---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse) 

data <- read_csv("Datacamp_Supermarket/loyalty.csv")

clean_data <- data %>%
	mutate(
		first_month = ifelse(first_month == ".", 0, first_month),
		region = factor(region, levels = c('Americas', 'Asia/Pacific', 'Europe', 'Middle East/Africa')),
		loyalty_years = factor(loyalty_years, order = TRUE, levels = c('0-1', '1-3', '3-5', '5-10', '10+')),
		joining_month = ifelse(is.na(joining_month) , "Unknown", joining_month),
		promotion = case_when(promotion == "YES" ~ "Yes", 
							  promotion == "NO" ~ "No", 
							 TRUE ~ promotion)
	) %>%
	mutate(
		first_month = as.numeric(first_month),
		joining_month = factor(joining_month, levels = c("Jan", "Feb", 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Unknown')), 
		promotion = factor(promotion, levels = c("No", "Yes"))
	)
clean_data

summary(clean_data)
levels(clean_data$joining_month)
```

# Task 2
The team at International Essentials have told you that they have always believed that the number of years in the loyalty scheme is the biggest driver of spend.

Producing a table showing the difference in the average spend by number of years in the loyalty programme along with the variance to investigate this question for the team.

You should start with the data in the file 'loyalty.csv'.
Your output should be a data frame named spend_by_years.
It should include the three columns loyalty_years, avg_spend, var_spend.
Your answers should be rounded to 2 decimal places.


```{r}
spend_by_years <- data %>%
  group_by(loyalty_years) %>% 
  summarize(
    avg_spend = round(mean(spend),2), 
    var_spend = round(var(spend),2)
  ) %>% 
  select(loyalty_years, avg_spend, var_spend) 

spend_by_years
```
Task 3
Fit a baseline model to predict the spend over the year for each customer.

Fit your model using the data contained in “train.csv”

Use “test.csv” to predict new values based on your model. You must return a dataframe named base_result, that includes customer_id and spend. The spend column must be your predicted 

```{r}
train <- read_csv("Datacamp_Supermarket/train.csv") %>% 
  mutate(
    loyalty_years = factor(loyalty_years, order = TRUE, levels = c('0-1', '1-3', '3-5', '5-10', '10+')))
    

model <- lm(spend~loyalty_years * first_month + items_in_first_month + joining_month, data = train)
summary(model)

test <- read_csv("Datacamp_Supermarket/test.csv") 


base_result <- test %>% 
  select(customer_id) %>% 
  mutate(spend = seq(1:n()))

base_result$spend <- predict(model, newdata = test)



# rmse <- round(sqrt(mean((train$spend - base_result$spend)^2)),3)
# 
# # Mostrar el resultado
# print(rmse)

```


Task 4
Fit a comparison model to predict the spend over the year for each customer.

Fit your model using the data contained in “train.csv”

Use “test.csv” to predict new values based on your model. You must return a dataframe named compare_result, that includes customer_id and spend. The spend column must be your predicted values.

```{r}
library(randomForest)
loyalty_levels <- c("0-1", "1-3", "3-5", "5-10", "10+")
train <- train %>%
  mutate(loyalty_encoded = as.numeric(factor(loyalty_years, levels = loyalty_levels, ordered = TRUE)))

test <- test %>%
  mutate(loyalty_encoded = as.numeric(factor(loyalty_years, levels = loyalty_levels, ordered = TRUE)))


features <- c("loyalty_encoded", "first_month", "items_in_first_month", "joining_month")

rf_model <- randomForest(x = train[, features], y = train$spend, ntree = 100)

predicted_spend <- predict(rf_model, newdata = test[, features])

compare_result <- test %>%
  select(customer_id) %>%
  mutate(spend = predicted_spend)
```

